---
title: "Homework 5"
subtitle: "ECON 470, Spring 2025"
author: "Ethan Murakami"
format:
  pdf:
    output-file: "murakami_e_hmwk5_s2"
    output-ext: "pdf"
    pdf-engine: xelatex
    tables: booktabs
    header-includes:
      - \usepackage{float}
      - \floatplacement{table}{H}
---


```{r}
#| include: false

if (!require("pacman")) install.packages("pacman")
pacman::p_load(tidyverse, ggplot2, dplyr, lubridate, readr, readxl, hrbrthemes, fixest,
               scales, gganimate, gapminder, gifski, png, tufte, plotly, OECD,
               ggrepel, survey, foreign, devtools, pdftools, kableExtra, modelsummary,
               kableExtra, broom, gridExtra)
```

Here is a link to my repository: {https://github.com/bemur3/hwk5}

\newpage 

```{r}
#| include: false
#| eval: true


load("/Users/ethanmurakami/Documents/GitHub/hwk5/submission2/Hwk5_workspace.RData")
```


\newpage 


## 1. Plot the share of the adult population with direct purchase health insurance over time.

```{r}
#| echo: false


print(direct_trend_plot)

```


\newpage 

## 2. Discuss the reduction in direct purchase health insurance in later years. Can you list a couple of policies that might have affected the success of the direct purchase insurance market?



\newpage 

## 3. Plot the share of the adult population with Medicaid over time

```{r}
#| echo: false

print(medicaid_trend_plot)

```



\newpage 

## 4. Plot the share of uninsured over time, separately by states that expanded Medicaid in 2014 versus those that did not. Drop all states that expanded after 2014.

```{r}
#| echo: false
print(uninsured.share.plot)


```


\newpage 

## 5. Calculate the average percent of uninsured individuals in 2012 and 2015, separately for expansion and non-expansion states. Present your results in a basic 2x2 DD table.

```{r}
#| echo: false

kable(dd_table_nice,
      digits = 2,
      booktabs = TRUE,
      caption = "DD Table for Medicaid Expansion",
      format = 'latex') %>%
  kable_styling(full_width = FALSE)
```

\newpage 

## 6. Estimate the effect of Medicaid expansion on the uninsurance rate using a standard DD regression estimator, again focusing only on states that expanded in 2014 versus those that never expanded.

```{r}
#| echo: false
modelsummary(dd_est,
             coef_rename = c("post" = "Post 2014", "treat" = "Expand", "post:treat" = "Post x Expand"),
             gof_omit = 'DF|F|Lik|AIC|BIC|Adj',
             title = "Standard DD Estimates for Medicaid Expansion")


```

\newpage

## 7. Include state and year fixed effects in your estimates. Try using the lfe or fixest package to estimate this instead of directly including the fixed effects.

```{r}
#| echo: false

modelsummary(models_twfe,
             coef_rename = c("post" = "Post 2014", "treat" = "Expand", "post:treat" = "Post x Expand"),
             gof_omit = 'DF|F|Lik|AIC|BIC|Adj',
             title = "DD Estimates: Standard vs TWFE")


```

\newpage

## 8. Repeat the analysis in question 7 but include all states (even those that expanded after 2014). Are your results different? If so, why?

```{r}
#| echo: false

modelsummary(models_dynamic,
             coef_rename = c("post" = "Post 2014", "treat" = "Expand",
                             "post:treat" = "Post x Expand", "treat_dynamic" = "Post x Expand (Dynamic)"),
             gof_omit = 'DF|F|Lik|AIC|BIC|Adj',
             title = "DD Estimates: Including Staggered Expansions")



```


\newpage

## 9. Provide an “event study” graph showing the effects of Medicaid expansion in each year. Use the specification that includes state and year fixed effects, limited to states that expanded in 2014 or never expanded.

```{r}
#| echo: false

iplot(mod_twfe,
      xlab = "Years Since Expansion",
      main = "Event Study (Common Treatment Year)")


```


\newpage

## 10.Repeat part 9 but again include states that expanded after 2014. Note: this is tricky…you need to put all states onto “event time” to create this graph.

```{r}
#| echo: false

iplot(mod_twfe2,
      xlab = "Years Since Treatment",
      main = "Event Study (Staggered Expansion Timing)") +
  geom_hline(yintercept = 0, linetype = "dashed", color = "black") +  # Horizontal line at 0
  geom_vline(xintercept = 0, linetype = "dashed", color = "red")       # Vertical line at event time 0

```